---
title: "A Cautionary Tale: The $TRUMP Token Case Study"
subtitle: "An Academic Analysis of Meme Coin Risk Factors"
author: "Minh Phuc Le (s3916144)"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: flatly
---

```{r setup, include=FALSE}
library(flexdashboard)
library(plotly)
library(dplyr)
library(tidyr)
library(DT)
library(lubridate)
library(scales)
library(readr)
library(zoo)
library(htmltools)

# Load the CSV data
trump_df <- read_csv("$trump price history.csv")
btc_df <- read_csv("btcusdmax.csv")
sol_df <- read_csv("solusdmax.csv")
holders_df <- read_csv("TOP$TRUMPtokenholders.csv")

# Data preprocessing with error handling
trump_df$date <- as.Date(trump_df$snapped_at)
btc_df$date <- as.Date(btc_df$snapped_at)
sol_df$date <- as.Date(sol_df$snapped_at)

# Filter Trump data from launch 
trump_clean <- trump_df %>% 
  filter(date >= as.Date("2025-01-17")) %>%
  arrange(date) %>%
  mutate(
    daily_return = (price - lag(price)) / lag(price) * 100,
    cumulative_return = ((price / first(price)) - 1) * 100,
    market_cap_billions = market_cap / 1e9,
    volume_millions = total_volume / 1e6,
    days_since_launch = as.numeric(date - min(date)) + 1
  ) %>%
  filter(!is.na(price), !is.na(total_volume))

# comparative data 
comparison_start <- min(trump_clean$date)
comparison_end <- max(trump_clean$date)

btc_period <- btc_df %>% 
  filter(date >= comparison_start, date <= comparison_end) %>%
  arrange(date) %>%
  mutate(
    daily_return = (price - lag(price)) / lag(price) * 100,
    cumulative_return = ((price / first(price)) - 1) * 100
  ) %>%
  filter(!is.na(price))

sol_period <- sol_df %>% 
  filter(date >= comparison_start, date <= comparison_end) %>%
  arrange(date) %>%
  mutate(
    daily_return = (price - lag(price)) / lag(price) * 100,
    cumulative_return = ((price / first(price)) - 1) * 100
  ) %>%
  filter(!is.na(price))

# Calculate key metrics
launch_price <- first(trump_clean$price)
current_price <- last(trump_clean$price)
peak_price <- max(trump_clean$price, na.rm = TRUE)
total_return <- ((current_price - launch_price) / launch_price) * 100
peak_return <- ((peak_price - launch_price) / launch_price) * 100
current_mcap <- last(trump_clean$market_cap_billions)
peak_mcap <- max(trump_clean$market_cap_billions, na.rm = TRUE)

# Volatility calculations
trump_vol <- sd(trump_clean$daily_return, na.rm = TRUE)
btc_vol <- sd(btc_period$daily_return, na.rm = TRUE)
sol_vol <- sd(sol_period$daily_return, na.rm = TRUE)

# Concentration metrics
top_holder_pct <- holders_df$Percentage[1]
top_10_pct <- sum(holders_df$Percentage[1:10], na.rm = TRUE)

# Risk events
extreme_days <- trump_clean %>%
  filter(abs(daily_return) > 20) %>%
  nrow()

total_days <- nrow(trump_clean)
```

# The Launch 

Column {data-width=400 data-height=200}
-----------------------------------------------------------------------

### **The Story Begins**

<div style="background-color: #d1ecf1; padding: 2px; border-radius: 10px; border-left: 5px solid #17a2b8; margin-bottom: 2px;">
  <h4 style="color: #0c5460; margin-bottom: 10px; font-size: 15.5px;"> Launch Price:</h4>
  <p style="font-size: 13.5px; color: #0c5460;">
    <strong>Price at Launch:</strong> $`r round(launch_price, 2)`<br>
  </p>
</div>

<div style="background-color: #d4edda; padding: 2px; border-radius: 10px; border-left: 5px solid #28a745; margin-bottom: 2px;">
  <h4 style="color: #155724; margin-bottom: 10px; font-size: 15.5px;"> Peak Price</h4>
  <p style="font-size: 13.5px; color: #155724;">
    <strong>Peak Price:</strong> $`r round(peak_price, 2)`<br>
  </p>
</div>

<div style="background-color: #f8d7da; padding: 2px; border-radius: 10px; border-left: 5px solid #dc3545;">
  <h4 style="color: #721c24; margin-bottom: 2px;font-size: 15.5px;"> Current Price</h4>
  <p style="font-size: 13.5px; color: #721c24;">
    <strong>Current Price:</strong> $`r round(current_price, 2)`<br>
    <strong>Total Return:</strong> `r ifelse(total_return > 0, "+", "")` `r round(total_return, 1)`% `  /  ` <strong>Days Since Launch:</strong> `r total_days` days
  </p>
</div>


### **The Initial Promise**

<div style="font-size: 12px; line-height: 1.4; color: #333;">
On January 17, 2025, the cryptocurrency community experienced something they have never heard of before, the appearance of a meme coin on the market related to Donald Trump, called the <strong>$TRUMP</strong> coin. This case study focuses on what followed next and the reason as to why even tokens associated with a president have excessive risks.<br>
  <strong>Key Launch Metrics:</strong><br>
  - Initial market capitalization: $`r round(first(trump_clean$market_cap_billions), 1)` billion<br>
  - Peak market cap reached: $`r round(peak_mcap, 1)` billion<br>
  - Time to peak: `r trump_clean$days_since_launch[which.max(trump_clean$price)]` days<br>
  <strong>The Academic Question:</strong><br>
  Does the essence of market govern the establishment of a sustainable cryptocurrency value despite the political prestige and social media influence.
</div>

Column {data-width=600}
-----------------------------------------------------------------------

### **The Price Journey: A Visual Timeline**

```{r}
# Enhanced price chart with narrative annotations
price_plot <- plot_ly(trump_clean, x = ~date, y = ~price, 
                     type = 'scatter', mode = 'lines',
                     line = list(color = '#2E86AB', width = 3),
                     name = '$TRUMP Price',
                     hovertemplate = paste0(
                       '<b>Date:</b> %{x}<br>',
                       '<b>Price:</b> $%{y:.2f}<br>',
                       '<b>Market Cap:</b> $%{customdata:.1f}B<br>',
                       '<extra></extra>'
                     ),
                     customdata = ~market_cap_billions) %>%
  layout(
    title = list(text = "$TRUMP Token: The Complete Price Story", 
                 font = list(size = 16, family = "Arial, sans-serif")),
    xaxis = list(title = "Date", 
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    yaxis = list(title = "Price (USD)", 
                 tickformat = "$,.2f",
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)',
    font = list(family = "Arial, sans-serif"),
    margin = list(t = 50, l = 60, r = 30, b = 50)
  )

# Add key event annotations
peak_date <- trump_clean$date[which.max(trump_clean$price)]
price_plot <- price_plot %>%
  add_annotations(
    x = peak_date, y = peak_price,
    text = paste0("Peak: $", round(peak_price, 2), "<br>", format(peak_date, "%b %d")),
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#A23B72",
    font = list(size = 10, color = "#A23B72"),
    bgcolor = "rgba(255,255,255,0.8)",
    bordercolor = "#A23B72", borderwidth = 1
  ) %>%
  add_annotations(
    x = first(trump_clean$date), y = launch_price,
    text = paste0("Launch: $", round(launch_price, 2)),
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#F18F01",
    font = list(size = 10, color = "#F18F01"),
    bgcolor = "rgba(255,255,255,0.8)",
    bordercolor = "#F18F01", borderwidth = 1
  )

price_plot
```

### **Volume Analysis: The Liquidity Story**

```{r}
# Volume analysis with academic insights
volume_plot <- plot_ly(trump_clean, x = ~date, y = ~volume_millions,
                      type = 'bar',
                      marker = list(color = '#C73E1D', opacity = 0.7),
                      name = 'Trading Volume',
                      hovertemplate = paste0(
                        '<b>Date:</b> %{x}<br>',
                        '<b>Volume:</b> $%{y:.1f}M<br>',
                        '<b>Price:</b> $%{customdata:.2f}<br>',
                        '<extra></extra>'
                      ),
                      customdata = ~price) %>%
  layout(
    title = list(text = "Daily Trading Volume: Liquidity Patterns", 
                 font = list(size = 16, family = "Arial, sans-serif")),
    xaxis = list(title = "Date", 
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    yaxis = list(title = "Volume (Millions USD)", 
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)',
    font = list(family = "Arial, sans-serif"),
    margin = list(t = 50, l = 60, r = 30, b = 50)
  )

volume_plot
```

# The Reality

Column {data-width=400}
-----------------------------------------------------------------------

### **Risk Assessment Framework**

<div style="background-color: #fff3cd; padding: 15px; border-radius: 10px; border-left: 5px solid #ffc107; font-size: 13px; color: #856404; margin-bottom: 20px; line-height: 1.6;">
  <h4 style="margin-top: 0; font-size: 14px;"> Academic Risk Metrics</h4>
  <ul style="margin: 0 0 10px 0; padding-left: 18px;">
    <li><strong>Herfindahl Index Equivalent:</strong> `r round((top_holder_pct/100)^2 + sum((holders_df$Percentage[-1]/100)^2, na.rm = TRUE), 3)`</li>
    <li><strong>Top 10 Concentration:</strong> `r round(top_10_pct, 1)`%</li>
    <li><strong>Gini Coefficient (est.):</strong> &gt; 0.85 (Extreme inequality)</li>
  </ul>

  <h4 style="margin-top: 0; font-size: 14px;"> Statistical Analysis</h4>
  <ul style="margin: 0; padding-left: 18px;">
    <li><strong>$TRUMP Volatility:</strong> `r round(trump_vol, 1)`% (σ)</li>
    <li><strong>Bitcoin Multiple:</strong> `r round(trump_vol/btc_vol, 1)`x more volatile</li>
    <li><strong>Extreme Days:</strong> `r extreme_days` out of `r total_days` (`r round(extreme_days/total_days*100, 1)`%)</li>
  </ul>
</div>



### **Volatility Comparison**

```{r}
# Academic volatility analysis
vol_data <- data.frame(
  Asset = c("$TRUMP", "Solana", "Bitcoin"),
  Volatility = c(trump_vol, sol_vol, btc_vol),
  Classification = c("Extreme", "High", "Moderate")
)

vol_colors <- c("Extreme" = "#C73E1D", "High" = "#F18F01", "Moderate" = "#2E86AB")

vol_plot <- plot_ly(vol_data, x = ~reorder(Asset, -Volatility), y = ~Volatility,
                   type = 'bar', 
                   color = ~Classification, colors = vol_colors,
                   text = ~paste0(round(Volatility, 1), "%"),
                   textposition = 'inside',
                   textfont = list(size = 12, color='white'),
                   hovertemplate = paste0(
                     '<b>%{x}</b><br>',
                     'Daily Volatility: %{y:.1f}%<br>',
                     'Classification: %{customdata}<br>',
                     '<extra></extra>'
                   ),
                   customdata = ~Classification) %>%
  layout(
    title = list(text = "Volatility Analysis (Daily Returns)", 
                 font = list(size = 14, family = "Arial, sans-serif")),
    xaxis = list(title = "", tickfont = list(size = 11)),
    yaxis = list(title = "Standard Deviation (%)", 
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    showlegend = FALSE,
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)',
    font = list(family = "Arial, sans-serif"),
    margin = list(t = 50, l = 60, r = 30, b = 50),
    height = 300
  )

vol_plot
```


Column {data-width=600}
-----------------------------------------------------------------------

### **Comparative Performance Analysis**

```{r}
# Create comparative returns chart
comparison_data <- data.frame(
  date = trump_clean$date,
  TRUMP = trump_clean$cumulative_return,
  Bitcoin = btc_period$cumulative_return[match(trump_clean$date, btc_period$date)],
  Solana = sol_period$cumulative_return[match(trump_clean$date, sol_period$date)]
) %>%
  pivot_longer(cols = -date, names_to = "Asset", values_to = "Return") %>%
  filter(!is.na(Return))

performance_plot <- plot_ly(comparison_data, x = ~date, y = ~Return, 
                            color = ~Asset, 
                            colors = c("TRUMP" = "#C73E1D", "Bitcoin" = "#F18F01", "Solana" = "#2E86AB"),
                            type = 'scatter', mode = 'lines',
                            line = list(width = 3),
                            hovertemplate = paste0(
                              '<b>%{x}</b><br>',
                              '<b>Asset:</b> %{data.name}<br>',
                              '<b>Return:</b> %{y:.1f}%<br>',
                              '<extra></extra>'
                            )) %>%
  layout(
    title = list(text = "Cumulative Returns: Comparative Analysis", 
                 font = list(size = 14, family = "Arial, sans-serif")),
    xaxis = list(title = "Date", 
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    yaxis = list(title = "Cumulative Return (%)", 
                 tickformat = ".1f",
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)',
    font = list(family = "Arial, sans-serif"),
    margin = list(t = 50, l = 60, r = 30, b = 80),
    legend = list(x = 0.98, y = 0.98),
        shapes = list(
      list(type = "line", 
           x0 = min(comparison_data$date),
           x1 = max(comparison_data$date),
           y0 = 0, y1 = 0,
           line = list(color = "gray", dash = "dash", width = 1))
    )
  )


performance_plot
```

### **Ownership Structure: The Concentration Problem**

```{r}
# Ownership concentration visualization
top_20 <- holders_df %>%
  slice(1:20) %>%
  mutate(
    Rank = row_number(),
    Cumulative = cumsum(Percentage)
  )

ownership_plot <- plot_ly() %>%
  add_trace(
    data = top_20,
    x = ~Rank, y = ~Cumulative,
    type = 'scatter', mode = 'lines+markers',
    line = list(color = '#C73E1D', width = 4),
    marker = list(color = '#C73E1D', size = 6),
    name = 'Actual Distribution',
    hovertemplate = paste0(
      '<b>Holder Rank:</b> %{x}<br>',
      '<b>Individual:</b> %{customdata:.1f}%<br>',
      '<b>Cumulative:</b> %{y:.1f}%<br>',
      '<extra></extra>'
    ),
    customdata = ~Percentage
  ) %>%
  add_trace(
    x = c(1, 20), y = c(5, 100),
    type = 'scatter', mode = 'lines',
    line = list(color = 'gray', dash = 'dash', width = 2),
    name = 'Equal Distribution',
    hoverinfo = 'skip'
  ) %>%
  layout(
    title = list(text = "Token Ownership Concentration Analysis", 
                 font = list(size = 14, family = "Arial, sans-serif")),
    xaxis = list(title = "Holder Rank", 
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    yaxis = list(title = "Cumulative Ownership (%)", 
                 tickfont = list(size = 11),
                 titlefont = list(size = 12)),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)',
    font = list(family = "Arial, sans-serif"),
    margin = list(t = 50, l = 60, r = 30, b = 50),
    legend = "bottomright"
  ) %>%
  add_annotations(
    x = 1, y = top_20$Percentage[1],
    text = paste0("Single holder:<br>", round(top_20$Percentage[1], 1), "%"),
    showarrow = TRUE, arrowhead = 2, arrowcolor = "#C73E1D",
    font = list(size = 10, color = "#C73E1D"),
    bgcolor = "rgba(255,255,255,0.8)",
    bordercolor = "#C73E1D"
  )

ownership_plot
```

# The Lesson 

Column {data-width=500 data-height=800}
-----------------------------------------------------------------------

### **Investment Scenario Analysis**

```{r}
# Investment scenarios table
scenarios <- data.frame(
  Investment = c("$1,000", "$5,000", "$10,000"),
  At_Peak = paste0(
    "$", format(round(c(1000, 5000, 10000) * (1 + peak_return / 100)), big.mark = ",")
  ),
  Current_Value = paste0(
    "$", format(round(c(1000, 5000, 10000) * (1 + total_return / 100)), big.mark = ",")
  ),
  Peak_Gain_Loss = paste0(
    ifelse(peak_return > 0, "+$", "-$"),
    format(round(abs(c(1000, 5000, 10000) * peak_return / 100)), big.mark = ",")
  ),
  Current_Gain_Loss = paste0(
    ifelse(total_return > 0, "+$", "-$"),
    format(round(abs(c(1000, 5000, 10000) * total_return / 100)), big.mark = ",")
  )
)

div(
  tags$style(HTML("
    table.dataTable thead th {
      font-size: 11px;
      text-align: center;
    }
  ")),
  DT::datatable(scenarios,
                options = list(dom = 't', pageLength = 10, scrollX = TRUE),
                colnames = c('Initial Investment', 'Peak Value', 'Current Value', 'Peak Gain/Loss', 'Current Gain/Loss'),
                caption = "Hypothetical Investment Scenarios",
                class = 'compact stripe') %>%
    DT::formatStyle(columns = 1:5, fontSize = '12px', textAlign = 'center')
)

```

<h4 style="margin-top: 20px;"> Academic Learning Framework</h4>

<div style="font-size: 13px; line-height: 1.5; color: #333; margin-top: 5px;">
<strong>Market Efficiency:</strong><br>
- Even with presidential association, fundamental market forces dominate causing voltality<br>
<strong>Risk-Return Profile:</strong><br>
- Extreme volatility (`r round(trump_vol, 1)`%) vs established crypto assets<br>
<strong>Behavioral Finance:</strong> <br>
- Initial euphoria followed by reality-based correction<br>
<strong>Quantitative Findings:</strong><br>
  - <strong>Concentration Risk:</strong> `r round(top_holder_pct, 1)`% single holder concentration<br>
  - <strong>Volatility Premium:</strong> `r round(trump_vol/btc_vol, 1)`x Bitcoin volatility<br>
  - <strong>Extreme Events:</strong> `r round(extreme_days/total_days*100, 1)`% of trading days<br>

  <strong>Regulatory Implications:</strong><br>
  - Celebrity/political tokens face unique regulatory scrutiny<br>
  - High concentration creates manipulation vulnerability<br>
  - Liquidity patterns suggest institutional absence
</div>


Column {data-width=550}
-----------------------------------------------------------------------

### **Academic Conclusions & Investment Framework**
<h4 style="margin-top: 0px;"> Empirical Evidence Summary:</h4>

<div style="font-size: 13px; line-height: 1.5; color: #333;">
  The <strong>$TRUMP</strong> token case study provides quantifiable evidence of extreme risk factors in celebrity/political meme coins:<br>
  <strong>Statistical Significance:</strong>
  - Volatility exceeds traditional crypto by factor of `r round(trump_vol/btc_vol, 1)`<br>
  <strong>Market Structure Risk:</strong>
  - Ownership concentration of `r round(top_holder_pct, 1)`% creates manipulation potential<br>
  <strong>Liquidity Analysis:</strong>
  - Volume inconsistency indicates institutional avoidance
</div>

**Investment Decision Framework:**

```{r}
# Investment suitability framework
suitability_data <- data.frame(
  Investor_Profile = c("Conservative", "Moderate", "Aggressive", "Speculative"),
  Tolerance = c("Low", "Medium", "High", "Very High"),
  Allocation = c("0%", "0%", "0.5-1%", "2-5%"),
  Suitability = c("Not Suitable", "Not Suitable", "Extreme Caution", "Consider"),
  Rationale = c(
    "Exceeds risk parameters",
    "Volatility incompatible", 
    "Only with stop-losses",
    "Speculation component only"
  )
)

div(
  tags$style(HTML("
    table.dataTable thead th {
      font-size: 11px;
      text-align: center;
    }
    .dataTables_wrapper {
      overflow-y: visible !important;
    }
  ")),
  
  DT::datatable(
    suitability_data,
    options = list(
      dom = 't',
      pageLength = 15,
      scrollX = TRUE,
      scrollY = FALSE  # Important: disables vertical scroll
    ),
    caption = "Investment Suitability by Risk Profile",
    class = 'compact stripe',
    rownames = FALSE
  ) %>%
    DT::formatStyle(
      'Suitability',
      backgroundColor = DT::styleEqual(
        c('Not Suitable', 'Extreme Caution', 'Consider'),
        c('#ffebee', '#fff3e0', '#e8f5e8')
      )
    ) %>%
    DT::formatStyle(
      columns = 1:5,
      fontSize = '11px',
      textAlign = 'center'
    )
)


```
<h4 style="margin-top: 10px;"> Final Assessment </h4>

<div style="font-size: 13px; line-height: 1.5; color: #333;">
The example of the <strong>$TRUMP</strong> token illustrates that even such an association with a presidential candidate is not enough to neutralize the basic risks of crypto-assets. Its excessive concentration, volatility, and the regulatory uncertainty associated with it cannot make it a part of any ordinary investment portfolio available; rather, it is a specimen to study speculative behavior of assets and market psychology in action.
</div>

**Data Sources:**

The data used for this dashboard was retrieved on CoinGecko and Solscan


